---
title: "Assignment"
author: "tawaku"
date: "2020/12/8"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Prediction for the Weight Lifting Exercise.

## Overview
In this assignment, we built the model to predict how well the excercise is done based on data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. 

## Analysis
Load libraries to be used in this analysis.
```{r load_libraries, message=FALSE, warning=FALSE, include=FALSE}
library(caret)
library(ggplot2)
library(dplyr)
library(corrplot)
library(tidyr)
library(visdat)
library(rpart)
library(rpart.plot)
library(randomForest)
```

Read dumbbell exercise data and replace invalid data to 'NA'.
```{r read_data}
data <- read.csv('data/pml-training.csv', na.strings = c('', 'NA', '#DIV/0!'))
quiz <- read.csv('data/pml-testing.csv', na.strings = c('', 'NA', '#DIV/0!'))
```

Find columns that have missing data. Remove columns whose 90% of data is missing.
```{r missing_columns}
set.seed(1234)
data %>% sample_n(size = 1000) %>% vis_miss()

na_count <- apply(data, 2, function(x){ sum(is.na(x)) })
label <- names(na_count[na_count < 0.9 * length(data)])

data <- data %>% select(any_of(label))
quiz <- quiz %>% select(any_of(label))
```

Remove columns that are not related to classe for prediction.
```{r remove_columns}
label <- label[!(label %in% 
  c(
    'X', 'user_name',
    'raw_timestamp_part_1', 'raw_timestamp_part_2', 
    'cvtd_timestamp', 'new_window', 'num_window'
  )
)]

data <- data %>% select(any_of(label))
quiz <- quiz %>% select(any_of(label))
```

Remove the pair of columns with high correlation to reduce the dimension.
```{r correlation}
cor_matrix <- cor(data[!names(data) == 'classe'])
corrplot(
  cor_matrix, 
  method = 'color', 
  type = 'lower', 
  order = 'hclust', 
  tl.cex = .90, 
  tl.col = 'black'
)
low_cor_index <- findCorrelation(cor_matrix, cutoff = .70, exact = TRUE)

data <- data %>% select(c(label[low_cor_index], 'classe'))
quiz <- quiz %>% select(label[low_cor_index])
```

Review cleaned data. 
```{r review_data}
summary(data)
```

Separate data to training/testing for cross validation.
```{r data_partition}
inTrain <- createDataPartition(y = data$classe, p = 0.75, list = FALSE)
training <- data[inTrain,]
testing <- data[-inTrain,]
```

Train the model with decision tree.
Calculate the accuracy of this model and it's 0.676.
```{r rpart}
rp_model <- rpart(classe ~ ., data = training, method = 'class')
rp_pred <- predict(rp_model, testing, type = 'class')

(rp_cm <- table(rp_pred, testing$classe))
sum(diag(rp_cm))/sum(rp_cm)

rpart.plot(rp_model)
```

Train the model with random forest.
Calculate the accuracy of this model and it's 0.987.
Based on the plot, important variables are roll_belt, yaw_belt and pitch_belt.
```{r random_forest}
rf_model <- randomForest(
    x = training[names(training) != 'classe'],
    y = training$classe,
    xtest = testing[names(testing) != 'classe'],
    ytest = testing$classe,
    keep.forest = TRUE
)
rf_model$confusion
sum(diag(rf_model$confusion)/sum(rf_model$confusion))
varImpPlot(rf_model)
```

Check if PCA will improve the accuracy of model. PC1 ~ PC11 explain about .975.
```{r pca}
pca_data <- prcomp(data[names(data) != 'classe'], center = TRUE, scale = TRUE)
summary(pca_data)
```

Built random forest with PCA data. PCA data doesn't improve the accuracy.
```{r train_with_pca_data}
pca_data <- data.frame(classe = data$classe, pca_data$x)
pca_data <- pca_data[,1:11]

inTrain <- createDataPartition(y = pca_data$classe, p = 0.75, list = FALSE)
pca_training <- pca_data[inTrain,]
pca_testing <- pca_data[-inTrain,]

pca_rf_model <- randomForest(
    x = pca_training[names(pca_training) != 'classe'],
    y = pca_training$classe,
    xtest = pca_testing[names(pca_testing) != 'classe'],
    ytest = pca_testing$classe,
    keep.forest = TRUE
)

pca_rf_model$confusion
sum(diag(pca_rf_model$confusion)/sum(pca_rf_model$confusion))
```

## Conclusion
RandomForest with raw dataset has highest accuracy among models we built to predict how well the excercise is done.
Predict quiz data with the model.
```{r predict}
predict(rf_model, newdata = quiz)
```

